\bf A Continuous-time Multicast \\Network Model

### 2018-05-09 paper/paper.tex
PPE results updated
added a paragraph motivating the out of sample experiment

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL 
\usepackage[algo2e]{algorithm2e}
\usepackage{algorithmic}  
\usepackage{algorithm}
\usepackage[toc,page]{appendix}
%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

@@

	Figure \ref{figure:PPCresults} illustrates the results of poterior predictive checks. Upper two plots show node-specific degree distributions, where the left one is the boxplots of 500 posterior predictive outdegree statistic and the right plot is the same one for indegree statistic. For both plots, the x-axis represents the node id's from 1 to 18, and the y-axis represents the number of emails sent or received by the node. When compared with the observed outdegree and indegree statistics (red lines), our model recovers the overall distribution of ties across the nodes. For example, node 1 and 10 show significantly higher level of both sending and receiving activities relative to the rest, and the model-simulated data captures those big jumps, without using node-specific indicators. Outdegree distribution of some low-activity nodes are not precisely recovered, however, indegree distribution looks much better. Since we use more information in the recipient selection process (i.e., network effects) while we only rely on minimum time-increments when choosing the observed sender, these results are expected. Lower left plot is the distribution of recipient sizes, where the x-axis spans over the size of recipients 1 to 14 (which is the maximum size of observed recipients) and the y-axis denotes the number of documents with x-number of recipients. The result shows that our model is underestimating single-receiver documents while overestimating documents with two, three, and four recipients. One explanation behind what we observe is that the model is trying to recover so-called ``broadcast'' emails, which are the documents with $\geq 10$ number of recipients, so that the intercept estimate $b_1$ is slightly moved toward right. To improve the goodness of fit for recipient sizes, we can add more covariates (e.g., sender indicators) or assign strong prior structure on $b_1$. In the end, the plot on the lower right is the PP plot for time-increments, which is a graphical measure commonly used for assessing how closely two data sets agree, by plotting the two cumulative distribution functions against each other. The resulting goodness of fit of the diagonal line gives a measure of the difference between the distribution of simulated time-increments and the observed ones, and our PP plot shows that we have great performance in reproducing the observed time distribution.
	\begin{figure}[!t]
		\includegraphics[width=0.495\textwidth]{plots_paper/outdegreenew-1.png}	
		\includegraphics[width=0.495\textwidth]{plots_paper/indegreenew-1.png}	
		\includegraphics[width=0.495\textwidth]{plots_paper/recipientsizenew-1.png}	
		\includegraphics[width=0.31\textwidth]{plots_paper/timeppnew.pdf}
		\caption {PPC results: outdegree distribution (\textit{upper left}), indegree distribution (\textit{upper right}), recipient size distribution (\textit{lower left}), and time-increments probability-probability (PP) plot  (\textit{lower right}). Red lines in the first three plot depict the observed statistics, and the red line in the last plot is the diagonal line connecting $(0, 0)$ and $(1, 1)$.}
		\label{figure:PPCresults}
	\subsubsection{Prediction Experiment}\label{subsubsec:Experiment_email}
Lastly, we evaluate the model's ability to predict ties from the Montgomery County email data, conditioned on their ``training" part of the data. To perform the experiment, we separately formed a test split of each three components---sender, recipients, and timestamps---by randomly selecting ``test" data with probability $p=0.1$. Any missing variables were imputed by drawing samples from their conditional posterior distributions, given the observed data, parameter estimates, and current values of test data. We then run inference to update the latent variables given the imputed and observed data. We iterate the two steps---imputation and inference---multiple times to obtain enough number of estimates for ``test" data. Algorithm \ref{alg:PPE} outlines this procedure in detail, including the full conditional distributions for missing sender, recciever, and timestamps.	
	\begin{algorithm}[!t]
		\caption{Out-of-Sample Tie Predictions}
		\label{alg:PPE}
		\begin{algorithmic}
			\STATE {\bfseries Input:} data $ \{ (a_d, \boldsymbol{r}_d, t_d)\}_{d=1}^D$, 
			number of new data to generate $R$,
			hyperparameters
			\vskip 0.1in
			\textbf{Test splits:}	
			\STATE draw test authors with $p=0.1$ (out of $D$ authors) 
			\STATE draw test recipients
			with $p=0.1$ (out of $D\times (A-1)$ recipient indicators $\{\{\boldsymbol{r}_{dr}\}_{r\in [A]_{\backslash a_d}}\}_{d=1}^D$)
			\STATE draw test timestamps with $p=0.1$  (out of $D$ timestamps) 
			\STATE set the ``test" data as ``missing" (NA)
			\vskip 0.1in
			\textbf{Imputation and inference:}	
			\STATE initialize the parameters $(\boldsymbol{b}, \boldsymbol{\eta}, \boldsymbol{u}, \sigma^2_\tau)$
			\FOR{$r=1$ {\bfseries to}  $R$}
			\FOR{$d=1$ {\bfseries to}  $D$}
			\IF{$a_d=$  NA}
			\FOR{$a=1$ {\bfseries to} $A$}
			\STATE compute $\pi_{a} $ using $P(a_d= a | \cdot)=f_{\tau}(\tau_{d}; \mu_{a_dd}, \sigma_\tau^2)\times \prod_{a\neq a_d}\big(1-F_{\tau}(\tau_{d}; \mu_{a d}, \sigma_\tau^2) \big)$
			\ENDFOR
			\STATE draw $a_d \sim \mbox{Multinomial}(\pi)$
			\ENDIF
			\FOR{$r\in [A]_{\backslash a_d}$}
			\IF{$r_{dr}=$ NA}
			\STATE draw $r_{dr}$ using $P(r_{dr}= 1 | \cdot)$ and $P(r_{dr}= 0| \cdot)$---Equation (7)
			\ENDIF
			\ENDFOR
			\IF{$t_d=$ NA}
			\STATE draw $\boldsymbol{\tau}^{new}_d$ from proposal distribution (e.g., $\boldsymbol{\tau}^{new}_d \sim N(\boldsymbol{\tau}^{old}_d, \sigma^2_{prop})$)
			\STATE use Metropolis-Hastings to decide accept or reject using the probability
			\begin{equation*}
			\footnotesize
			\frac{f_{\tau}(\tau^{new}_{d}; \mu_{a_d d}, \sigma_\tau^2)\times \prod_{a\neq a_d}\big(1-F_{\tau}(\tau_{d}; \mu_{a d}, \sigma_\tau^2) \big) }{f_{\tau}(\tau^{old}_{d}; \mu_{a_d d}, \sigma_\tau^2)\times \prod_{a\neq a_d}\big(1-F_{\tau}(\tau_{d}; \mu_{a d}, \sigma_\tau^2) \big)},
			\end{equation*}
			where $\boldsymbol{\tau}^{old}_d$ is from earlier iteration.
			\ENDIF
			\STATE run inference and update $(\boldsymbol{b}, \boldsymbol{\eta},\boldsymbol{u}, \sigma^2_\tau)$ given the imputed and observed data
			\ENDFOR
			\STATE store the estimates for ``test" data
			\ENDFOR
		\end{algorithmic}

@@

After running the experiment, we compare the prediction accuracy

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL 
\usepackage[algo2e]{algorithm2e}
\usepackage{algorithmic}  
\usepackage{algorithm}
\usepackage[toc,page]{appendix}
%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

@@

	Figure \ref{figure:PPCresults} illustrates the results of poterior predictive checks. Upper two plots show node-specific degree distributions, where the left one is the boxplots of 500 posterior predictive outdegree statistic and the right plot is the same one for indegree statistic. For both plots, the x-axis represents the node id's from 1 to 18, and the y-axis represents the number of emails sent or received by the node. When compared with the observed outdegree and indegree statistics (red lines), our model recovers the overall distribution of ties across the nodes. For example, node 1 and 10 show significantly higher level of both sending and receiving activities relative to the rest, and the model-simulated data captures those big jumps, without using node-specific indicators. Outdegree distribution of some low-activity nodes are not precisely recovered, however, indegree distribution looks much better. Since we use more information in the recipient selection process (i.e., network effects) while we only rely on minimum time-increments when choosing the observed sender, these results are expected. Lower left plot is the distribution of recipient sizes, where the x-axis spans over the size of recipients 1 to 14 (which is the maximum size of observed recipients) and the y-axis denotes the number of documents with x-number of recipients. The result shows that our model is underestimating single-receiver documents while overestimating documents with two, three, and four recipients. One explanation behind what we observe is that the model is trying to recover so-called ``broadcast'' emails, which are the documents with $\geq 10$ number of recipients, so that the intercept estimate $b_1$ is slightly moved toward right. To improve the goodness of fit for recipient sizes, we can add more covariates (e.g., sender indicators) or assign strong prior structure on $b_1$. In the end, the plot on the lower right is the PP plot for time-increments, which is a graphical measure commonly used for assessing how closely two data sets agree, by plotting the two cumulative distribution functions against each other. The resulting goodness of fit of the diagonal line gives a measure of the difference between the distribution of simulated time-increments and the observed ones, and our PP plot shows that we have great performance in reproducing the observed time distribution.
	\begin{figure}[!t]
		\centering
		\includegraphics[width=0.495\textwidth]{plots_paper/outdegreenew-1.png}	
		\includegraphics[width=0.495\textwidth]{plots_paper/indegreenew-1.png}	
		\includegraphics[width=0.495\textwidth]{plots_paper/recipientsizenew-1.png}	
		\includegraphics[width=0.31\textwidth]{plots_paper/timeppnew.pdf}
		\caption {PPC results: outdegree distribution (\textit{upper left}), indegree distribution (\textit{upper right}), recipient size distribution (\textit{lower left}), and time-increments probability-probability (PP) plot  (\textit{lower right}). Red lines in the first three plot depict the observed statistics, and the red line in the last plot is the diagonal line connecting $(0, 0)$ and $(1, 1)$.}
		\label{figure:PPCresults}
	\end{figure}
	\subsubsection{Prediction Experiment}\label{subsubsec:Experiment_email}

The HeEM ({\bf Hyper-edge event model)	has many component parts that need to be specified by the user (i.e., the selection of the event timing features ($\bm{y}$), the recipient selection features $(\bm{x}$), and the event time distribution ($g$)). Many of these components will be specified based on user expertise (e.g., regarding which features would drive recipient selection), but some decisions may require a data-driven approach to model specification. For example, though theoretical considerations may inform the specification of features, subject-matter expertise is unlikely to inform the decision regarding the family of the event time distribution. Furthermore, since different distribution families (and model specifications more generally) may involve different size parameter spaces, any data-driven approach to model comparison must guard against over-fitting the data. In this section we present a general-purpose approach to evaluating a HeEM specification using out-of-sample prediction. We illustrate this approach by comparing alternative distributional families for the event timing component of the model.

Lastly, we evaluate the model's ability to predict ties from the Montgomery County email data, conditioned on their ``training" part of the data. To perform the experiment, we separately formed a test split of each three components---sender, recipients, and timestamps---by randomly selecting ``test" data with probability $p=0.1$. Any missing variables were imputed by drawing samples from their conditional posterior distributions, given the observed data, parameter estimates, and current values of test data. We then run inference to update the latent variables given the imputed and observed data. We iterate the two steps---imputation and inference---multiple times to obtain enough number of estimates for ``test" data. Algorithm \ref{alg:PPE} outlines this procedure in detail, including the full conditional distributions for missing sender, recciever, and timestamps.	
	\begin{algorithm}[!t]
		\spacingset{1}
		\caption{Out-of-Sample Tie Predictions}
		\label{alg:PPE}
		\begin{algorithmic}
			\STATE {\bfseries Input:} data $ \{ (a_d, \boldsymbol{r}_d, t_d)\}_{d=1}^D$, 
			number of new data to generate $R$,
			hyperparameters
			\vskip 0.1in
			\textbf{Test splits:}	
			\STATE draw test authors with $p=0.1$ (out of $D$ authors) 
			\STATE draw test recipients
			with $p=0.1$ (out of $D\times (A-1)$ recipient indicators $\{\{\boldsymbol{r}_{dr}\}_{r\in [A]_{\backslash a_d}}\}_{d=1}^D$)
			\STATE draw test timestamps with $p=0.1$  (out of $D$ timestamps) 
			\STATE set the ``test" data as ``missing" (NA)
			\vskip 0.1in
			\textbf{Imputation and inference:}	
			\STATE initialize the parameters $(\boldsymbol{b}, \boldsymbol{\eta}, \boldsymbol{u}, \sigma^2_\tau)$
			\FOR{$r=1$ {\bfseries to}  $R$}
			\FOR{$d=1$ {\bfseries to}  $D$}
			\IF{$a_d=$  NA}
			\FOR{$a=1$ {\bfseries to} $A$}
			\STATE compute $\pi_{a} $ using $P(a_d= a | \cdot)=f_{\tau}(\tau_{d}; \mu_{a_dd}, \sigma_\tau^2)\times \prod_{a\neq a_d}\big(1-F_{\tau}(\tau_{d}; \mu_{a d}, \sigma_\tau^2) \big)$
			\ENDFOR
			\STATE draw $a_d \sim \mbox{Multinomial}(\pi)$
			\ENDIF
			\FOR{$r\in [A]_{\backslash a_d}$}
			\IF{$r_{dr}=$ NA}
			\STATE draw $r_{dr}$ using $P(r_{dr}= 1 | \cdot)$ and $P(r_{dr}= 0| \cdot)$---Equation (7)
			\ENDIF
			\ENDFOR
			\IF{$t_d=$ NA}
			\STATE draw $\boldsymbol{\tau}^{new}_d$ from proposal distribution (e.g., $\boldsymbol{\tau}^{new}_d \sim N(\boldsymbol{\tau}^{old}_d, \sigma^2_{prop})$)
			\STATE use Metropolis-Hastings to decide accept or reject using the probability
			\begin{equation*}
			\footnotesize
			\frac{f_{\tau}(\tau^{new}_{d}; \mu_{a_d d}, \sigma_\tau^2)\times \prod_{a\neq a_d}\big(1-F_{\tau}(\tau_{d}; \mu_{a d}, \sigma_\tau^2) \big) }{f_{\tau}(\tau^{old}_{d}; \mu_{a_d d}, \sigma_\tau^2)\times \prod_{a\neq a_d}\big(1-F_{\tau}(\tau_{d}; \mu_{a d}, \sigma_\tau^2) \big)},
			\end{equation*}
			where $\boldsymbol{\tau}^{old}_d$ is from earlier iteration.
			\ENDIF
			\STATE run inference and update $(\boldsymbol{b}, \boldsymbol{\eta},\boldsymbol{u}, \sigma^2_\tau)$ given the imputed and observed data
			\ENDFOR
			\STATE store the estimates for ``test" data
			\ENDFOR
		\end{algorithmic}
	\end{algorithm}

@@

After running the experiment, we measure the predictive performance of two separate time distributions using 500 predicted samples, which are depicted in Figure \ref{figure:PPEresults}.
	\begin{figure}[!b]
		\centering
		\includegraphics[width=0.33\textwidth]{plots_paper/outdegreenew-1.png}	
		\includegraphics[width=0.33\textwidth]{plots_paper/indegreenew-1.png}	
		\includegraphics[width=0.33\textwidth]{plots_paper/recipientsizenew-1.png}	
		\caption {PPE results}
		\label{figure:PPEresults}
	\end{figure}
